The complex network of trophic interactions in a subAntarctic oceanic Marine Protected Area

### 2024-01-11 MS_ToSubmit/OriginallySubmittedVersion.Rmd
Fix format

---
title: The complex network of trophic interactions in a subAntarctic oceanic Marine Protected Area
author:
  - name: Tomás I. Marina^a^
    email: tomasimarina@gmail.com
    correspondingauthor: true
  - name: Irene R. Schloss^a^^,^^b^^,^^c^
  - name: Gustavo A. Lovrich^a^
  - name: Claudia C. Boy^a^
  - name: Daniel O. Bruno^a^^,^^c^
  - name: Fabiana L. Capitanio^d^^,^^e^
  - name: Sergio M. Delpiani^f^
  - name: Juan Martín Díaz de Astarloa^f^
  - name: Cintia Fraysse^a^
  - name: Virginia A. García Alonso^d^^,^^e^
  - name: Andrea Raya Rey^a^^,^^c^^,^^g^
  - name: Laura Schejter^h^
  - name: Mariela L. Spinelli^d^^,^^e^
  - name: Marcos Tatián^i^^,^^j^
  - name: Diego Urteaga^k^
  - name: Luciana Riccialdelli^a^

address:
  - code: 1
    address: Centro Austral de Investigaciones Científicas (CADIC-CONICET), Ushuaia, Argentina
  - code: 2
    address: Instituto Antártico Argentino (IAA), CABA, Argentina
  - code: 3
    address: Instituto de Ciencias Polares, Ambiente y Recursos Naturales (ICPA), Universidad de Tierra del Fuego (UNTDF), Ushuaia, Argentina
  - code: 4
    address: Facultad de Ciencias Exactas y Naturales, Universidad de Buenos Aires (UBA), CABA, Argentina
  - code: 5
    address: Instituto de Biodiversidad y Biología Experimental y Aplicada (IBBEA), Universidad de Buenos Aires-CONICET, CABA, Argentina
  - code: 6
    address: Instituto de Investigaciones Marinas y Costeras (IIMYC), Universidad Nacional de Mar del Plata-CONICET, Mar del Plata, Argentina
  - code: 7
    address: Wildlife Conservation Society, Argentina
  - code: 8
    address: Instituto Nacional de Investigación y Desarrollo Pesquero (INIDEP), Consejo Nacional de Investigaciones Científicas y Técnicas (CONICET), Mar del Plata, Argentina
  - code: 9
    address: Facultad de Ciencias Exactas, Físicas y Naturales, Universidad Nacional de Córdoba (UNC), Córdoba, Argentina
  - code: 10
    address: Instituto de Diversidad y Ecología Animal (IDEA-CONICET), Córdoba, Argentina
  - code: 11
    address: Museo Argentino de Ciencias Naturales “Bernardino Rivadavia”, CABA, Argentina
abstract: |
  Abstract: The total area of the world ocean designated under marine protection has increased recently. Most Marine Protected Areas (MPAs) target vulnerable, keystone, charismatic and or endemic species. In the sub-Antarctic, ocean protection is associated to oceanic islands,except for MPAs Namuncurá - Burdwood Bank I and II (MPA N-BB, ~53º–55ºS and ~56º–62ºW), which are associated to a submarine plateau and its southern adjacent deep slope, respectively. Here, we present the first analysis of the network of predator-prey interactions for the MPA N-BB. We applied a network approach to characterise the complexity and structure of the food web, and identify the species' role in such a framework. The MPA N-BB food web consisted of 1788 interactions and 379 species, with a connectance of 0.01. Almost half of the consumers were omnivores (0.48), and the network displayed a small-world pattern. These suggest that the ecosystem might be vulnerable to perturbations targeting highly connected species, although other properties might provide resilience and resistance, resulting in a rearranged structure that preserves its original functions. Several species arose as important in terms of different aspects of trophic structure and functioning, and response to perturbations. Generalist species, mainly fishes, play a crucial role in the ecosystem's benthopelagic coupling and should be considered as relevant energy transfers for the ecosystem. We argue that the diversity of species, including both the benthic and pelagic habitats, is responsible for securing the connectivity within the food web against perturbations, therefore contributing to the structure and stability of the ecosystem.
keywords: 
  - Food Web
  - Complexity
  - Structure
  - Burdwood Bank
  - Southwest Atlantic
classoption: preprint, 3p, authoryear
bibliography: AMPBancoBurdwood.bib
linenumbers: true
numbersections: true
# Use a CSL with `citation_package = "default"`
csl: "elsevier-harvard2.csl"
output: 
  rticles::elsevier_article:
    latex_engine: xelatex
    keep_tex: true
    citation_package: "default"
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

library(knitr)
library(tidyverse)
library(pander)
library(ggplot2)
library(kableExtra)
library(multiweb)
library(ggjoy)
library(ggpubr)
library(igraph)
library(gridExtra)
library(grid)

# Introduction

@@

# Methodology

## Study area

@@

```{r figure1, echo=FALSE, fig.width = 5, fig.height = 10, out.width = '100%', fig.cap="Marine Protected Areas Namuncurá - Burdwood Bank I (MNR and MNP, northern section) and II (MNR and SMNR, southern section). Acronyms indicate categories according to the management plan: MNR - Marine National Reserve, MNP - Marine National Park and RMNR - Restricted Marine National Reserve."}

## Network construction

@@

## Network analysis

@@

```{r table1, echo=FALSE, results='asis'}

@@

# Results

## Network-level properties

```{r figure2, echo=FALSE, fig.width = 5.5, fig.height = 10, fig.align='center', fig.cap="Graph of the food web for the MPA Namuncurá - Burdwood Bank. Circles represent species and arrows trophic interactions. Circle diameter is relative to the number of interactions. Colour gradient indicates the trophic level."}
# Food web graph

@@

@@

```{r table2, echo=FALSE}

knitr::kable(round(prop_topol,2) %>% dplyr::select(Size, Links, LD, Connectance, Omnivory, PathLength, Clustering) %>% mutate(SW = "True"), col.names = c("Species","Interactions","Density","Connectance","Omnivory","CPL", "CC", "SW"),
                      caption = "Network-level properties of the MPA Namuncurá - Burdwood Bank food web. CPL: Characteristic Path Length; CC: Clustering Coefficient; SW: Small-World pattern. See table 1 for definitions and ecological relevance.") %>% row_spec(0,bold=TRUE)

```{r figure3, echo=FALSE, fig.width = 6.5, fig.height = 9, fig.align='center', fig.cap="Degree distributions for the (A) food web, for (B) prey among predators, (C) predators among prey, and (D) for each functional group. Groups are vertically ordered by increasing trophic level (following coloration of figure 2); groups with less than 3 species were not plotted (e.g., pelagic fish). All functional groups and the species that comprise them are shown in Supplementary Material (Table S3)."}
# Degree distributions: network, predator, prey & group
# above row (network, prey, predator), below row (group)

@@

# Move to a new page
grid.newpage()
# Nested ggarrange
figure_3 <- ggarrange(a,                                                                                         # First row
          ggarrange(b, c, ncol = 2, labels = c("B", "C"), font.label=list(color="black",size=10)),   # Second row
          ggarrange(d, ncol = 1),                                                                    # Third row
          nrow = 3, labels = "A",
          font.label=list(color="black",size=10))
figure_3

## Species-level properties

```{r figure4, echo=FALSE, fig.width = 6.5, fig.height = 6.5, fig.align='center', fig.cap="Species-level properties by trophic level: (A) betweenness, (B) closeness, (C) trophic similarity, and (D) topological role. Each point represents a species. Linear regressions for betweenness ($y = 74.97x - 117.35, R^2 = 0.05, p-value < 0.01$), closeness ($y = 9.33e-06x - 9.31e-4, R^2 = 0.003, p-value = 0.15$) and trophic similarity ($y = -0.02x + 0.11 , R^2 = 0.07, p-value < 0.01$). Note that for panels A, B and C only species with TLs equal or greater than 2 were considered."}
# Betweenness, Closeness, Trophic similarity & Topological role by Trophic level

@@

# Move to a new page
grid.newpage()
# Arrange figure
figure_ab <- ggarrange(plot_sp_ts + rremove("xlab"), plot_sp_cl + rremove("xlab"),
                      labels = c("A","B"), ncol=2,
                      font.label = list(size = 10, color = "black"))
figure_ab <- annotate_figure(figure_ab, bottom = textGrob("Trophic level", gp = gpar(fontface="bold", col="black", fontsize=10)))
figure_4 <- ggarrange(figure_ab,                                                                                         # First row
          ggarrange(plot_sp_bt, plot_sp_tr, ncol = 2, labels = c("C", "D"), font.label=list(color="black",size=10)),                                            
          nrow = 2)
figure_4

# Discussion

## The food web of the MPA Namuncurá - Burdwood Bank ecosystem

## Dominant consumers and food sources

## Species' role related to their trophic level

## Caveats and future perspectives

# Conclusion

# Acknowledgements

We are indebted to all those experts of the working group 'Banco Burdwood' who humbly provided their knowledge to enhance the quality of the present research. Although most of them are authors of the present work, it is worth to mention the following researchers: Brenda L. Doti (IBBEA, CONICET-UBA; Universidad de Buenos Aires, Argentina), Sofía L. Callá (Museo Argentino de Ciencias Naturales “Bernardino Rivadavia”, Argentina), Sandra Gordillo (IDACOR-CONICET; Universidad Nacional de Córdoba, Argentina), Mariano I. Martinez (Museo Argentino de Ciencias Naturales “Bernardino Rivadavia”, Argentina) and Luciano Padovani (Instituto Nacional de Investigación y Desarrollo Pesquero, INIDEP, Argentina). We thank the MPA Namuncurá − Burdwood Bank administration. Research cruises were funded by national funds by the Law 26.875. This study was funded by Consejo Nacional de Investigaciones Científicas y Técnicas (CONICET) and Agencia Nacional de Promoción Científica y Tecnológica (PICT 2020-SERIEA-01617), Argentina. This work is contribution no. XX of the MPA Namuncurá (Law 26.875).

# Literature Cited

>>>>

---
output:
  pdf_document:
    latex_engine: xelatex
    keep_tex: yes
  word_document: default
bibliography: AMPBancoBurdwood.bib
csl: "elsevier-harvard2.csl"
documentclass: article
geometry:
- vmargin=1in
- hmargin=1in
header-includes:
- \usepackage{amsmath}
- \usepackage{pdflscape,booktabs}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
- \usepackage[running]{lineno}
- \linenumbers
---

# The complex network of trophic interactions in a subAntarctic oceanic Marine Protected Area

## Abstract

The total area of the world ocean designated under marine protection has increased recently. Most Marine Protected Areas (MPAs) target vulnerable, keystone, charismatic and or endemic species. In the sub-Antarctic, ocean protection is associated to oceanic islands,except for MPAs Namuncurá - Burdwood Bank I and II (MPA N-BB, ~53º–55ºS and ~56º–62ºW), which are associated to a submarine plateau and its southern adjacent deep slope, respectively. Here, we present the first analysis of the network of predator-prey interactions for the MPA N-BB. We applied a network approach to characterise the complexity and structure of the food web, and identify the species' role in such a framework. The MPA N-BB food web consisted of 1788 interactions and 379 species, with a connectance of 0.01. Almost half of the consumers were omnivores (0.48), and the network displayed a small-world pattern. These suggest that the ecosystem might be vulnerable to perturbations targeting highly connected species, although other properties might provide resilience and resistance, resulting in a rearranged structure that preserves its original functions. Several species arose as important in terms of different aspects of trophic structure and functioning, and response to perturbations. Generalist species, mainly fishes, play a crucial role in the ecosystem's benthopelagic coupling and should be considered as relevant energy transfers for the ecosystem. We argue that the diversity of species, including both the benthic and pelagic habitats, is responsible for securing the connectivity within the food web against perturbations, therefore contributing to the structure and stability of the ecosystem.

Keywords: Food Web, Complexity, Structure, Burdwood Bank, Southwest Atlantic

## 1. Introduction

@@

## 2. Methodology

### 2.1. Study area

@@

### 2.2. Network construction

@@

### 2.3. Network analysis

@@

## 3. Results

### 3.1. Network-level properties

### 3.2. Species-level properties

## 4. Discussion

### 4.1. The food web of the MPA Namuncurá - Burdwood Bank ecosystem

### 4.2. Dominant consumers and food sources

### 4.3. Species' role related to their trophic level

### 4.4. Caveats and future perspectives

## 5. Conclusion

## Acknowledgements

We are indebted to all those experts of the working group 'Banco Burdwood' who humbly provided their knowledge to enhance the quality of the present research. Although most of them are authors of the present work, it is worth to mention the following researchers: Brenda L. Doti (IBBEA, CONICET-UBA; Universidad de Buenos Aires, Argentina), Sofía L. Callá (Museo Argentino de Ciencias Naturales “Bernardino Rivadavia”, Argentina), Sandra Gordillo (IDACOR-CONICET; Universidad Nacional de Córdoba, Argentina), Mariano I. Martinez (Museo Argentino de Ciencias Naturales “Bernardino Rivadavia”, Argentina) and Luciano Padovani (Instituto Nacional de Investigación y Desarrollo Pesquero, INIDEP, Argentina). We thank the MPA Namuncurá - Burdwood Bank administration. Research cruises were funded by national funds by the Law 26.875. This study was funded by Consejo Nacional de Investigaciones Científicas y Técnicas (CONICET) and Agencia Nacional de Promoción Científica y Tecnológica (PICT 2020 SERIEA 01617), Argentina. This work is contribution no. XX of the MPA Namuncurá (Law 26.875).

## Tables

```{r table1, echo=FALSE, message=FALSE, results='asis'}

require(tidyverse)
require(pander)

@@

```{r tableS2, echo=FALSE, message=FALSE, results='asis'}
require(pander)
require(tidyverse)

@@

pandoc.table(round(prop_topol,2) %>% dplyr::select(Size, Links, LD, Connectance, Omnivory, PathLength, Clustering) %>% mutate(SW = "True"), col.names = c("Species","Interactions","Density","Connectance","Omnivory","CPL", "CC", "SW"), split.cells = c(1,1,1,1,1,1,1,1),
                      caption = "Network-level properties of the MPA Namuncurá - Burdwood Bank food web. CPL: Characteristic Path Length; CC: Clustering Coefficient; SW: Small-World pattern. See table 1 for definitions and ecological relevance.")

## Figures

```{r figure1, echo=FALSE, message=FALSE, warning=FALSE, fig.width = 5, fig.height = 10, out.width = '100%', fig.cap="Marine Protected Areas Namuncurá - Burdwood Bank I (MNR and MNP, northern section) and II (MNR and SMNR, southern section). Acronyms indicate categories according to the management plan: MNR - Marine National Reserve, MNP - Marine National Park and RMNR - Restricted Marine National Reserve."}

```{r figure2, echo=FALSE, message=FALSE, warning=FALSE, fig.width = 5.5, fig.height = 10, fig.align='center', fig.cap="Graph of the food web for the MPA Namuncurá - Burdwood Bank. Circles represent species and arrows trophic interactions. Circle diameter is relative to the number of interactions. Colour gradient indicates the trophic level."}
# Food web graph

# Packages
require(multiweb)

# Data
load("../results/summary_results_jul23.rda")

@@

```{r figure3, echo=FALSE, message=FALSE, warning=FALSE, fig.width = 6.5, fig.height = 9, fig.align='center', fig.cap="Degree distributions for the (A) food web, for (B) prey among predators, (C) predators among prey, and (D) for each functional group. Groups are vertically ordered by increasing trophic level (following coloration of figure 2); groups with less than 3 species were not plotted (e.g., pelagic fish). All functional groups and the species that comprise them are shown in Supplementary Material (Table S3)."}
# Degree distributions: network, predator, prey & group
# above row (network, prey, predator), below row (group)

# Packages
require(tidyverse)
require(igraph)
require(grid)
require(ggpubr)
require(ggplot2)
require(ggjoy)

@@

# Move to a new page
grid.newpage()
# Nested ggarrange
figure_3 <- ggarrange(a,                                                                                         # First row
          ggarrange(b, c, ncol = 2, labels = c("B", "C"), font.label=list(color="black",size=10)),   # Second row
          ggarrange(d, ncol = 1),                                                                    # Third row
          nrow = 3, labels = "A",
          font.label=list(color="black",size=10))
figure_3
```

```{r figure4, echo=FALSE, message=FALSE, warning=FALSE, fig.width = 6.5, fig.height = 6.5, fig.align='center', fig.cap="Species-level properties by trophic level: (A) trophic similarity, (B) closeness, (C) betweenness, and (D) topological role. Each point represents a species. Linear regressions for betweenness ($y = 74.97x - 117.35, R^2 = 0.05, p-value < 0.01$), closeness ($y = 9.33e-06x - 9.31e-4, R^2 = 0.003, p-value = 0.15$) and trophic similarity ($y = -0.02x + 0.11 , R^2 = 0.07, p-value < 0.01$). Note that for panels A, B and C only species with TLs equal or greater than 2 were considered."}
# Betweenness, Closeness, Trophic similarity & Topological role by Trophic level
require(tidyverse)
require(igraph)
require(grid)
require(ggpubr)

@@

# Move to a new page
grid.newpage()
# Arrange figure
figure_ab <- ggarrange(plot_sp_bt + rremove("xlab"), plot_sp_cl + rremove("xlab"),
                      labels = c("A","B"), ncol=2,
                      font.label = list(size = 10, color = "black"))
figure_ab <- annotate_figure(figure_ab, bottom = textGrob("Trophic level", gp = gpar(fontface="bold", col="black", fontsize=10)))
figure_4 <- ggarrange(figure_ab,                                                                                         # First row
          ggarrange(plot_sp_ts, plot_sp_tr, ncol = 2, labels = c("C", "D"), font.label=list(color="black",size=10)),                                            
          nrow = 2)
figure_4
```

## References
